---
title: "Exercise_09"
output: html_document
date: "2024-03-18"
---
```{r}
library("tidyverse")
library("dplyr")
library("mosaic")
library("ggplot2")
library("infer")
library("skimr")
library("cowplot")
library("broom")
```
Step 1

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Street_et_al_2017.csv"
d <- read_csv(f)
fivenum <- skim(d)[c(2,10:16)] %>%
  na.omit() %>%
  rename(
    mean = numeric.mean,
    sd = numeric.sd,
    minimum = numeric.p0,
    fqtl = numeric.p25,
    median = numeric.p50,
    thqtl = numeric.p75,
    maximum = numeric.p100
    )
str(fivenum)
```
The dataframe fivenum summarizes the mean, sd and crucial five numbers for all numerical variables.


Step 2
```{r}
#par(mfrow=c(2,2))

p1 <- ggplot(data = d, aes(x = Group_size, y = ECV)) + geom_point()
p2 <- ggplot(data = d, aes(x = Longevity, y = ECV)) + geom_point()
p3 <- ggplot(data = d, aes(x = Weaning, y = ECV)) + geom_point()
p4 <- ggplot(data = d, aes(x = Repro_lifespan, y = ECV)) + geom_point()

plot_grid(p1,p2,p3,p4)
```
Step 3
```{r}
d3 <- d %>%
  drop_na(ECV, Group_size)
```


```{r}
d3 <- mutate(d3, centered_ECV = ECV - mean(ECV))
d3 <- mutate(d3, centered_Group_size = Group_size - mean(Group_size))
```

In the following chunks, two coefficients are calculated and named by themselves.
```{r}
ECV <- d3$ECV
GS <- d3$Group_size

(beta1 <- cor(ECV, GS) * (sd(ECV)/sd(GS)))
```

```{r}

(beta0 <- mean(ECV) - beta1 * mean(GS))
```

Step 4
```{r}
m <- lm(ECV ~ Group_size, data = d3)
m$coefficients

```

The intercept equals the beta0 calculated by hand; the slope equals beta1 calculated by hand.

Step 5
```{r}
lm_radiation <- function(radiation) {
temp <- d3 %>%
  filter(Taxonomic_group == radiation)
ECV <- temp$ECV
GS <- temp$Group_size
beta1 <- cor(ECV, GS) * (sd(ECV)/sd(GS))
beta0 <- mean(ECV) - beta1 * mean(GS)
print(beta0)
print(beta1)
}
```

In the following three chunks, the outputs are the intercept and the slope for each radiation respectively.
```{r}
betas_ct <- lm_radiation("Catarrhini")
```

```{r}
betas_pl <- lm_radiation("Platyrrhini")
```


```{r}
betas_st <- lm_radiation("Strepsirhini")
```

Beta0s (intercepts) are highly different; those of platyrrhines and strepsirhines share relatively closer intercepts. Platyrrhines and strepsirhines share relatively similar beta1s (slopes), whereas they are different from that of catarrhines.

Step 6
```{r}
d6 <- d3 %>%
  mutate(ECVhat = beta0 + beta1 * Group_size) %>%
  mutate(residuals = ECV -ECVhat)
str(d6)
```

SE for the slope coefficient by hand:
```{r}
(df_regression <- 1) 
(df_error <- nrow(d6) - df_regression - 1) 
```

```{r}
SSE <- sum((d6$residuals)^2)
SSE
```

```{r}
(MSE <- SSE/df_error) 
```

```{r}
SSX <- sum((d6$Group_size - mean(d6$Group_size))^2)
SEbeta1 <- sqrt(MSE/SSX)
SEbeta1

```

SE for the slope coefficient extracted from the model:
```{r}
m.summary <- tidy(m)
tidy(m)$std.error[2]
tidy(m)$p.value[2]
```
In the chunk above, the p value is extracted as well.

The 95% CI by hand:
```{r}
alpha <- 0.05
```

```{r}
(lower <- beta1 - qt(1 - alpha/2, df = nrow(d6) - 2) * SEbeta1)
(upper <- beta1 + qt(1 - alpha/2, df = nrow(d6) - 2) * SEbeta1)
```

The 95% CI extracted from model:
```{r}
CI <- confint(m, level = 1 - alpha)
CI[2,1:2]
```

p value calculated by hand:
```{r}
calc.statistic <- (beta1 - 0)/SEbeta1
(p <- 2 * pt(calc.statistic, df = nrow(d6) - 2, lower.tail = FALSE))
```
p value was already extracted earlier.

Step 8
We need to permute slope coefficient estimates. The sentence from Module 18 explain what we need to permute: "essentially, each permutation involves randomly shuffling values of the response variable under the null hypothesis that they are independent of the explanatory variable."
```{r}
alpha <- 0.05
confidence_level <- 1 - alpha
p_lower <- alpha/2
p_upper <- 1 - (alpha/2)
degrees_of_freedom <- nrow(d6) - 2
critical_value <- qt(p_upper, df = degrees_of_freedom)

original.slope <- m %>%
tidy(conf.int = TRUE, conf.level = confidence_level) %>%
mutate(lower = estimate - std.error * critical_value, upper = estimate + std.error *
    critical_value) %>%
    filter(term == "Group_size")
original.slope 
```

```{r}
permuted.slope <- d6 %>%
  specify(ECV ~ Group_size) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope")

head(permuted.slope)
```

```{r}
permuted.slope.summary <- permuted.slope %>%
  summarize(
    estimate = mean(stat)
  )

permuted.slope.summary
```

```{r}
(p.value <- permuted.slope %>%
  get_p_value(obs_stat = original.slope$estimate, direction="two_sided"))
```
The p value is 0.

Step 8
```{r}
boot.slope <- d6 %>%
  specify(ECV ~ Group_size) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "slope")

head(boot.slope)
```
```{r}
boot.slope.summary <- boot.slope %>%
  summarize(
    estimate = mean(stat)
  )

boot.slope.summary
```

```{r}
(CI.percentile <- get_ci(boot.slope, level = 1 - alpha, type = "percentile"))
(CI.theory <- get_ci(boot.slope, level = 1 - alpha, type = "se", point_estimate = pull(boot.slope.summary,
    estimate)))
```

These CIs suggest that beta1 (slope coefficient) is different from zero.
